\documentclass[11pt]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{enumerate}
\usepackage{fullpage}
\begin{document}
\title{LAB 1}
\author{Jack Sullivan and Ari Kobren}
\maketitle

\section{Olympics Main Components}
Our message passing system is built as a hierarchy. We use a top-level
\emph{Olympics} class for organization.  This class doesn't hold any
information about teams or scores but establishes the system in which
all messages are passed.  In particular, the system established by the
\emph{Olympics} class includes 5 child \emph{actors}: a
\emph{TeamRoster}, an \emph{EventRoster}, a
\emph{TabletRequestRouter}, a \emph{CacofonixListener} and an
\emph{EventSubscription}.

\subsection{TeamRoster}
The \emph{TeamRoster} stores a list of teams that are participating
in the olympic games.  When the \emph{TeamRoster} recieves a
\emph{TeamMessage} it looks up the corresponding \emph{Team} and
routes the message accordingly.  The \emph{TeamRoster} instructs that
\emph{Team} able to send any reply message to the initial sender of
the \emph{TeamMessage}.

\subsection{EventRoster}
The \emph{EventRoster} is very similar to the \emph{TeamRoster}: the
\emph{EventRoster} stores a list of the events taking place at the
olympics and is able to communicate with a number \emph{Event}
classes.  Wwhen it receives an \emph{EventMessage}, it routes the
message to the corresponding \emph{Event}.  It also instructs that
\emph{Event} to send any reply to the initial sender of the
\emph{EventMessage}.

\subsection{TabletRequestRouter}
The \emph{TabletRequestRouter} serves as a router that helps spread
client request to various workers.  At initialization, the
\emph{TabletRequestRouter} spawns 5 \emph{TabletRequestWorker}s to
which it routes the messages that it receives.  These workers are
asynchronous and allow our system to handle more simultaneous
requests.  Each \emph{TabletReequestWorker} is able to communicate
with the \emph{EventRoster} and \emph{TeamRoster}. If one of the
\emph{TabletRequestWorkers} dies, the \emph{TabletRequestRouter} will
spawn a new one in its place. The \emph{TabletRequestRouter} routes
messages to the workers using round-robin-scheduling.

\subsection{CacofonixListener}
\emph{CacofonixListener} is very similar to a
\emph{TabletRequestWorker}: it also stores is able to communicate with
the \emph{TeamRoster} and the \emph{EventRoster}.  The difference
between the \emph{CacofonixListener} and a \emph{TabletRequestWorker}
is that only a special class--\emph{CacofonixClient}--is able to
communicate with the \emph{CacofonixListener}.

\subsection{EventSubscription}
Each event has a corresponding \emph{EventSubscription} class. An
\emph{EventSubscription} class is responsible for keeping track of who
has subscribed to that event. When an \emph{EventSubscription} class
receives an \emph{EventScore} update message, it sends corresponding
update messages to each of its subscribers.  When it receives a
\emph{Subscription} message, it stores the sender in its list of
subscribers.

\section{System Runtime Workflow}
To begin running our system, we run the \emph{Olympics} class.  As
described above, this class spawns the \emph{TeamRoster},
\emph{EventRoster}, \emph{TabletRequestRouter},
\emph{CacofonixListener} and \emph{EventSubscription} actors.  Once
the system is in place, we start $n$ \emph{TabletClients}.  These
clients are equipped with the \emph{getScore}, \emph{getMedalTally}
and \emph{registerClient} methods. When a client calls its
\emph{getScore} or \emph{getMedalTally} methods a corresponding
\emph{EventMessage} containing an event name and a \emph{GetScore}
object or a \emph{TeamMessage} containing a team name and a
\emph{GetMedalTally} object is sent to the \emph{TabletRequestRouter}
(which is then forwarded appropriately).  Upon receiving a message, each
client prints the corresponding message to the screen.  When calling
the \emph{registerClient} method, a \emph{Subscribe} message is
generated for a particular event and then sent to the
\emph{Subscription} actor.  From then on, whenever there are updates
made to that event, an update will be sent back to the client.  In
this way, our system can accomodate a client-pull and server-push
architectures.

In addition to the $n$ \emph{TabletClients}, we also spawn a
\emph{CacofonixClient} actor.  This client is equipped with the
\emph{setScore} and \emph{incrementMedalTally} methods.  Calling
either of these methods generates either an \emph{EventMessage}
containing an event name and a \emph{SetEventScore} object or a
\emph{TeamMessage} containing a team name and an
\emph{IncrementMedals} object.  Messages from the
\emph{CacofonixClient} are always sent to the \emph{CacofonixListener}
as described above.

\section{Design, Bottlenecks and Potential Improvements}
We tried separate concerns as much as possible so that implementation
and debugging would be simplest.  As such, each our actors only
handles a few message types (and we a few, extremely simple
messge types).  We also wanted to make our system as
asynchronous as possible so we made sure to include the
\emph{TabletRequestRouter} and \emph{TableRequestWorkers} which can
process messages in parallel.

The main bottleneck we see in our system are our \emph{TeamRoster} and
\emph{EventRoster} classes.  This is because, all messages having to
do with any team or any event has to go through one of these classes.
We could improve our system by making both the \emph{TeamRoster} and
\emph{EventRoster} into routers (special Akka classes) so that they
could handle messages in parallel.  Additionally, we'd need to
implement a custom routing strategy that would look at the content of
the message and route the message appropriately (e.g. if handling a
\emph{GetScore} message for event $A$, the \emph{EventRoster} would
route the message to the event class representing event $A$).

We could also improve our system further by making it more fault
tolerant.  Instead of simple watching the \emph{TabletRequestWorkers}
and respawning them upon failure, we could watch classes like the
\emph{TabletRequestRouters}, events and teams.

\section{Results}
In our first experiment, we ran our code using 1, 5 and 10 clients
each with a rate of one request per .5 seconds and measured the min,
max and average response times.

\begin{tabular}{c|c|c|c}
  NUM CLIENTS & MIN & MAX & AVG \\
  \hline
  1  & 0.01s & 0.02s & 0.01385s \\
  5  & 0.01s & 0.03s & 0.01342s \\
  10 & 0.01s & 0.12s & 0.01475s \\
\end{tabular}

In our second experiment we ran our code using 5 clients with .01, .1
and 0.4 seconds between requests and measured the min, max and
average response times.


\begin{tabular}{c|c|c|c}
  REQUEST FREQ. & MIN & MAX & AVG \\
  \hline
  0.01s & 0.01s & 0.3s  & 0.11667s \\
  0.1s  & 0.01s & 0.07s & 0.01266s \\
  0.4s  & 0.01s & 0.06s & 0.01369s \\
\end{tabular}

As we observe, adding more clients seems to increase average
latency.  However, there is no real difference between having 1 client
and 5 clients.  We hypothesize that adding a tremendous number of
clients would have a serious effect on the latency.

Also, we observe that having the clients make requests at higher
frequencies severely affects the latency.  Again, there is not much
difference between a request rate of 0.4s and 0.1s however, when we
increase the request rate by an order of magnitude we see that the
average latency per request increases substantially.

\section{Software}
We've built our system on top of the Akka \emph{actors} library.  This
library provides a hierarchical message passing architecture for the
Scala programming language.

\section{How to Run}

\end{document}
